---
title: "Solubility Prediction Step 1"
author: "Mahitab Sakr"
format:
  gfm:
    echo: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Importing libraries: 

```{r}
library(tidyverse)
library(patchwork) 
```

Reading the data file;
```{r}
data <- read.csv("../data/curated-solubility-dataset-5.csv", 
                        sep = ",", header = TRUE)
```
Seeing the structure of the columns in the data file:
```{r}
str(data, vec.len=2)
```
Assigning the columns to be removed into a vector called ID_info:
```{r}
ID_info <- c("ID","Name", "InChI", "InChIKey", "SMILES", "SD", "Ocurrences")
```
Assigning the columns of the data file to a new file without the ID_info:
```{r}
df = data[,!(names(data) %in% ID_info)]
```
Seeing the structure of the columns in the new data file:
```{r}
str(df, vec.len=2)
```
Adding a new column called Group_num:
```{r}
df$Group_num <- NA
```
Assigning the values based on the numbers in the column "Group".
```{r}
df$Group_num[df$Group == "G1"] = 1
df$Group_num[df$Group == "G2"] = 2
df$Group_num[df$Group == "G3"] = 3
df$Group_num[df$Group == "G4"] = 4
df$Group_num[df$Group == "G5"] = 5
```
Looking at the resulted column values:
```{r}
table(df$Group_num)
```
Removing the old Group column from the data file:
```{r}
df = subset(df, select = -c(Group))
```
Loop to create a histogram for each column in the data file: (To check for normal distribution)
```{r}
#| label: Histograms
plots <- lapply(names(df), function(column) { 
  ggplot(df, aes(x = .data[[column]])) + 
    geom_histogram(bins = 30, fill = "#560178", color = "white") + 
    labs(title = column, x = column, y = NULL) + 
    theme_minimal(base_size = 9) + 
theme(plot.title = element_text(size = 10, face = "bold")) 
}) 
plots
```
Creating a new column called Lipinski:
```{r}
df$Lipinski <- NA
```
Assigning the values based on the Lipinski rule of 5.
```{r}
df$Lipinski[df$MolWt < 500 & df$MolLogP < 5 & df$NumHAcceptors < 10 & df$NumHDonors < 5] = 0
df$Lipinski[df$MolWt < 500 & df$MolLogP > 5 & df$NumHAcceptors < 10 & df$NumHDonors < 5] = 1
df$Lipinski[df$MolWt < 500 & df$MolLogP < 5 & df$NumHAcceptors > 10 & df$NumHDonors < 5] = 1
df$Lipinski[df$MolWt < 500 & df$MolLogP < 5 & df$NumHAcceptors < 10 & df$NumHDonors > 5] = 1
df$Lipinski[df$MolWt > 500 & df$MolLogP < 5 & df$NumHAcceptors < 10 & df$NumHDonors < 5] = 1
df$Lipinski[df$MolWt > 500 & df$MolLogP > 5 & df$NumHAcceptors < 10 & df$NumHDonors < 5] = 2
df$Lipinski[df$MolWt > 500 & df$MolLogP < 5 & df$NumHAcceptors > 10 & df$NumHDonors < 5] = 2
df$Lipinski[df$MolWt > 500 & df$MolLogP < 5 & df$NumHAcceptors < 10 & df$NumHDonors > 5] = 2
df$Lipinski[df$MolWt < 500 & df$MolLogP > 5 & df$NumHAcceptors > 10 & df$NumHDonors < 5] = 2
df$Lipinski[df$MolWt < 500 & df$MolLogP > 5 & df$NumHAcceptors < 10 & df$NumHDonors > 5] = 2
df$Lipinski[df$MolWt < 500 & df$MolLogP < 5 & df$NumHAcceptors > 10 & df$NumHDonors > 5] = 2
df$Lipinski[df$MolWt > 500 & df$MolLogP > 5 & df$NumHAcceptors > 10 & df$NumHDonors < 5] = 3
df$Lipinski[df$MolWt > 500 & df$MolLogP < 5 & df$NumHAcceptors > 10 & df$NumHDonors > 5] = 3
df$Lipinski[df$MolWt < 500 & df$MolLogP > 5 & df$NumHAcceptors > 10 & df$NumHDonors > 5] = 3
df$Lipinski[df$MolWt > 500 & df$MolLogP > 5 & df$NumHAcceptors > 10 & df$NumHDonors > 5] = 4
```
Looking at the resulted column values:
```{r}
table(df$Lipinski)
```
Writting it to take this version of the data file to python 
```{r}
write.csv(df, "df.csv")
```